Renewal and stability in populations structured by remaining years of life

### 2014-08-27 ThanoRepro/ThanoRepro.tex
.

\usepackage[top=2in, bottom=1.5in, left=1in, right=1in]{geometry}

% matrix numbering
\usepackage{trivfloat}
\trivfloat{matrix}
\floatstyle{plaintop}
    \restylefloat{matrix}
\AtBeginDocument{\numberwithin{matrix}{section}}
% end preamble
%-------------------------------------------------------

@@

All demographic forces vary over age in known and regular ways. Information on
such forces, along with the size and age structure of a population allows the
demographer to make predictions about the future size and structure of a population. In this
paper, I explore some formal demographic consequences of a particular
redefinition of age. Instead of counting age as the time passed between birth and the present (or
some other moment), consider age as the amount of time left from the present
until death\footnote{This quantity is referred to as \textit{residual} life in
the reliability literaure.}.
Individuals in this case move in the same direction along imaginary life lines, but the reference point for age is now at the end of the
life line instead of at the beginning. Of course, such information cannot be
requested of survey respondents,
because one's future time of death is typically unknown. It would be
possible to organize lifecourse data by remaining years of life in a
very long running linked population register, but such registers are few and far
between, and the analysis would be very retrospective in nature.
Instead we approximate this perspective using a strategy similar to that used by
\citet{brouard1986structure}, \citet{miller2001increasing}, or
\citet{lee2002approach}, which uses lifetables and aggregated population counts to redistribute counts within age-classes over
future death times. This allows us to make statements about the remaining
lifelines within a current stock of population, under certain assumptions.

>>>>

\usepackage[top=2in, bottom=1.5in, left=1in, right=1in]{geometry}
\usepackage{setspace}
% matrix numbering
\usepackage{trivfloat}
\trivfloat{matrix}
\floatstyle{plaintop}
    \restylefloat{matrix}
\AtBeginDocument{\numberwithin{matrix}{section}}
% end preamble
%-------------------------------------------------------

@@

\onehalfspacing
All demographic forces vary over age in known and regular ways. Information on
such forces, along with the size and age structure of a population allows the
demographer to make predictions about the future size and structure of a population. In this
paper, I explore some formal demographic consequences of a particular
redefinition of age. Instead of counting age as the time passed between birth and the present (or
some other moment), consider age as the amount of time left from the present
until death\footnote{This quantity is referred to as \textit{residual} lifetime
in the reliability literaure.}. Individuals in this case move in the same
direction along imaginary life lines, but the reference point for age is now at
the end of the life line instead of at the beginning. For individuals, the
timing of events in life is mirrored exactly between the two definitions of age,
but aggregate patterns differ.

Direct information on the timing of demographic transitions with respect to
death is only available from long-running panel or linked historical data
studies that are few and far between. Some surveys also ask respondents in various ways about
their perceived future lifetime. Perceived remaining lifetime may underly those
demographic transitions governed to some degree by volition, but this is not
treated here. The present inquiry is related to projected
remaining lifetime, using lifetable methods, which is an approximation of
actual remaining lifetime.
The lifetable methods applied here follow those of \citet{brouard1986structure}, \citet{miller2001increasing}, which redistribute age-classified counts over future
death times according to a fixed mortality schedule. This allows us to make
statements about the remaining lifelines within a current stock of population, under certain assumptions.
